Gut-resident microorganisms and their genes are associated with cognition and neuroanatomy in children

### 2025-01-06 manuscript/main.tex

\usepackage{scicite}
% Clean citations with biblatex
\usepackage[
backend=biber,
style=numeric,
maxbibnames=100
]{biblatex}

@@

Members of the RESONANCE consortium also contributed to
project planning, conceptualization, and funding acquisition.

@@

All data needed to evaluate the conclusions in the paper
are present in the paper and/or the Supplementary Materials.
Taxonomic and functional microbial profiles, as well as subject
demographics necessary for statistical analyses and machine learning are
available on the Data Dryad
\cite{bonhamDataDryad2023}.
Data processing, generation of summary statistics, and
generation of plots was performed using the julia programming language
and 3rd party packages
\cite{bezansonJuliaFreshApproach2017,bonhamMicrobiomeJlBiobakeryUtils2021,danischMakieJlFlexible2021,blaomMLJJuliaPackage2020,ben_sadeghi_2022_7359268,dahua_lin_2023_7695673,douglas_bates_2023_7734970}.
Information for replicating the package environment
as well as code for data analysis and figure generation, as well as scripts for automated
download of input files are available on GitHub at https://github.com/Klepac-Ceraj-Lab/Resonance,
and archived on Zenodo \cite{kevin_bonham_2023_7647510}.

@@

\begin{figure}[!htb]
    \centering
    \small
    % \includegraphics[width=0.9\textwidth]{assets/Figure3.png}
    \caption{
        \textbf{Random forest models predict both concurrent and future measured cognitive function.}
        Comparison of RF predictor importance versus linear models for children
        over 18 months (\textbf{A}).
        Colors represent whether the species belong to the group of
        top-important features that account for 60\% of the cumulative
        importance on the RF model, if that species was significant (q
        \textless{} 0.2) in linear models, both, or neither.
        Comparison of RF predictor importance for microbial features
        between concurrent and future cognitive function measure
        (\textbf{B}).
        Colors represent whether the species belong to the group of
        top-important features that account for 60\% of the cumulative
        importance on the RF model only for concurrent prediction,
        only for the future prediction, or for both models.
        Features absent of one model are shown on the origin-aligned sideplots.
        Distribution of ages for each microbiome sampling - cognitive assesment dyad (\textbf{C}).
        Blue violin plot shows distribution of sampling age,
        while red violin plot shows distribution of asessment age.
        Black lines connect ages for each individual subject.      
        Ranked feature importance for taxa in RF models for
        both concurrent (\textbf{D}) and future (\textbf{E})
        cognitive function measurement prediction.
        Red dots indicate how relative importances change (among taxa)
        when demographics (sex, SES) are arred as covariables.
        Abundances and prevalences of taxa that are important
        for RF models both for concurrent (\textbf{F}) and future (\textbf{G}) assesment prediction.
        }
    \label{fig:3}
\end{figure}

>>>>

\usepackage{scicite}
% Clean citations with biblatex
\usepackage[
backend=biber,
style=numeric,
maxbibnames=50
]{biblatex}

@@

@@

All data needed to evaluate the conclusions in the paper
are present in the paper and/or the Supplementary Materials.
Taxonomic and functional microbial profiles, as well as subject
demographics necessary for statistical analyses and machine learning are
available on the Data Dryad
\cite{bonhamDataDryad2023}.
Data processing, generation of summary statistics, and
generation of plots was performed using the julia programming language
and 3rd party packages
\cite{bezansonJuliaFreshApproach2017,bonhamMicrobiomeJlBiobakeryUtils2021,danischMakieJlFlexible2021,blaomMLJJuliaPackage2020,ben_sadeghi_2022_7359268,dahua_lin_2023_7695673,douglas_bates_2023_7734970}.
Information for replicating the package environment
as well as code for data analysis and figure generation, as well as scripts for automated
download of input files are available on GitHub at https://github.com/KlepacCerajLab/Resonance,
and archived on Zenodo \cite{kevin_bonham_2023_7647510}.

@@

\begin{figure}[!htb]
    \centering
    \small
    % \includegraphics[width=0.9\textwidth]{assets/Figure3.png}
    \caption{
        \textbf{Random forest models predict both concurrent and future measured cognitive function.}
        (\textbf{A}) Comparison of RF predictor importance versus linear models for children
        over 18 months.
        Colors represent whether the species belong to the group of
        top-important features that account for 60\% of the cumulative
        importance on the RF model, if that species was significant (q
        \textless{} 0.2) in linear models, both, or neither.
        Ranked feature importance for taxa in RF models for
        both concurrent (\textbf{B}) and future (\textbf{C})
        cognitive function measurement prediction.
        Red dots indicate how relative importances change (among taxa)
        when demographics (sex, SES) are arred as covariables.
        (\textbf{D}) Comparison of RF predictor importance for microbial features
        between concurrent and future cognitive function measure.
        Colors represent whether the species belong to the group of
        top-important features that account for 60\% of the cumulative
        importance on the RF model only for concurrent prediction,
        only for the future prediction, or for both models.
        Features absent of one model are shown on the origin-aligned sideplots.
        Abundances and prevalences of taxa that are important
        for RF models both for concurrent (\textbf{E}) and future (\textbf{F}) assesment prediction.
        (\textbf{G}) Distribution of ages for each microbiome sampling - cognitive assesment dyad
        at sampling age (left),
        or asessment age (right).
        Black lines connect ages for each individual subject.
        }
    \label{fig:3}
\end{figure}
